A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-11-26 paper/paper.tex
paper: minor pluralization
paper: minor polish

This new construction is useful in the direction of better understanding the technicality of coherence issues in defining ``semi-simplicial types''~\citep{part15,shulman14,altenkirch16,kraus17}, as well as the dialectic between reasoning in Extensional Type Theory and Intensional Type Theory (as will be discussed in Section~\ref{sec:ett}).

@@

% 5. The "mathematical" construction (and its intuition)
%   leaving equality and inequality side conditions implicit
%   in particular, this means:
%   - assuming equality reflection
%   - equality as in observational type theory
%   - assuming termination of recursive constructions
% 6. Formal details
% - making substitution of equals by equals explicit
% - observational properties of equality
% - the theory of inequality proofs
% - the well-foundedness
% - the formalization tricks
%   - implicit arguments
%   - refine-style vs tactic-style

@@

Sections~\ref{sec:wf}, \ref{sec:le}, and \ref{sec:eqproperties} describe additional issues to be addressed in order to get a fully formal construction:
\begin{enumerate}
  \item The well-foundedness of the induction requires a special termination evidence which will be discussed in section~\ref{sec:wf}.
  \item The construction is indexed over integers and holds under some constraints on the range of these integers. There is a standard formalization dilemma in this kind of situation: either the constraints on the range are embedded in the construction so that the construction makes sense only on the corresponding range, or the construction is made first on a more general domain than needed but restricted to a smaller domain at the time of use. We adopted the first approach, requiring however the construction to be dependent on proofs of inequalities on natural numbers. We discuss how we deal with such dependencies in section~\ref{sec:le}.
  \item A number of standard groupoid properties of equality as well as type isomorphisms have been left implicit in the informal definition. This is discussed in section~\ref{sec:eqproperties}.
\end{enumerate}

@@

\section{Future work}
The construction could be extended with degeneracies as well as with permutations~\citep{grandis03}. Dependent $\nu$-sets could also be defined, opening the way to construct $\Pi$-types and $\Sigma$-types of $\nu$-sets. A $\nu$-set of $\nu$-set representing a universe could also be defined (as sketched e.g. in a talk by Herbelin at the HoTT-UF workshop for the bridge case~(\citeyear{herbelin-hott-uf})). More generally, we believe these lines of work to eventually provide alternative models to parametric type theories~\citep{nuyts17,cavallo19}.

>>>>

This new construction is useful in the direction of better understanding the technicality of coherence issues in defining ``semi-simplicial types''~\citep{part15,shulman14,altenkirch16,kraus17}, as well as the dialectic between reasoning in Extensional Type Theory and Intensional Type Theory.

@@

@@

Sections~\ref{sec:wf}, \ref{sec:le}, and \ref{sec:eqproperties} describe additional issues to be addressed in order to get a fully formal construction:
\begin{enumerate}
  \item The well-foundedness of the induction requires a special termination evidence which will be discussed in section~\ref{sec:wf}.
  \item The construction is indexed over integers and holds under some constraints on the range of these integers. There is a standard formalization dilemma in this kind of situation: either the constraints on the range are embedded in the construction so that the construction makes sense only on the corresponding range, or the construction is made first on a more general domain than needed but restricted to a smaller domain at the time of use. We adopted the former approach, requiring the construction to be dependent on proofs of inequalities on natural numbers. We discuss how we deal with such dependencies in section~\ref{sec:le}.
  \item A number of standard groupoid properties of equality as well as type isomorphisms have been left implicit in the informal definition. This is discussed in section~\ref{sec:eqproperties}.
\end{enumerate}

@@

\section{Future work}
The construction could be extended with degeneracies as well as with permutations~\citep{grandis03}. Dependent $\nu$-sets could also be defined, opening the way to construct $\Pi$-types and $\Sigma$-types of $\nu$-sets. A $\nu$-set of $\nu$-sets representing a universe could also be defined (as sketched e.g. in a talk by Herbelin at the HoTT-UF workshop for the bridge case~(\citeyear{herbelin-hott-uf})). More generally, we believe these lines of work to eventually provide alternative models to parametric type theories~\citep{nuyts17,cavallo19}.
